Lies and the lying liars who tell them

### 2018-11-02 GSV-comment-brief.Rmd
Cleaner pipeline when you rerun the java; use new continuous Bayesian updating over lambda, and use population sampling approach

library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE)
options(huxtable.knitr_output_format = "latex")
source("bayesian-heads.R")
suppressPackageStartupMessages({
  library(Hmisc)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(tidyr)
  library(readr)
  library(ggplot2)
  library(broom)
  library(huxtable)
})

@@

nreps <- 1000L

# set to FALSE to recalculate all results; TRUE for speed:
cache_results <- FALSE

# set to TRUE to (re)create the simulations CSV file:
rewrite_csv <- FALSE

# if TRUE, the number of liars in each simulation is randomly drawn
# from a population with `Liars` liars.
# If FALSE, the number of liars is fixed at `Liars * N`.
# If you change this, you may need to recreate the simulations CSV file
sample_from_pop <- FALSE

@@

TODO XXX: point out the claims they make. Add your sims of when the posterior is wrong
and intervals fail even for (R-T)/(N-T).

@@

Some people are honest, while others will lie whenever it benefits them. We would like to understand
the prevalence of lying, because dishonesty may be economically and socially harmful. Since we
cannot simply ask people if they are liars, one way to estimate the proportion of liars in a group
is to ask them to report the result of a coin flip or other random device, offering them a payment
if they report heads. Liars don't always lie: they only lie when it benefits them. So they always
report heads irrespective of the true coin flip. ^[GSV maintain this assumption and so shall I.] If
there are many more heads than we would expect by chance, we can assume many people are lying. But
how many?

@@

  path_root <- paste0("GSV-sims", if (sample_from_pop) "-pop-sample")

if (rewrite_csv) {
  readr::write_csv(sim_output, path = paste0(path_root, ".csv"))

  # the java only lets you select a document from your home directory
  copied_to_home_dir_ok <- file.copy(path, normalizePath("~"), overwrite = TRUE)
  stopifnot(copied_to_home_dir_ok)
  # now run java -jar LyingCalculator.jar from the LyingCalculator directory
  # make sure the old output file (xxxOutput.csv) is deleted first, or it will be appended to!
}

@@

output_path <- paste0(path_root, "Output.csv")
gsv_estimates <- suppressMessages(readr::read_csv(normalizePath(paste0("~/", output_path))))

@@

my_prior <- uniform_prior
# alternative:
# my_prior <- function (N) dbinom(0:N, N, p = 0.5)

bayes_est <- sim_output %>% 
      rowwise() %>% 
      do({
        pstr <- update_prior(
                heads = .$R, 
                N = .$N, 
                P = .$P/100, 
                prior = my_prior(.$N)
              )
        ci <- .$CI / 100
        ci_qs <- c(1/2 - ci/2, 1/2 + ci/2)
        cis <- posterior_quantile(ci_qs, pstr)
        # calculate quantities as prop dishonest
        tibble(
          EV.ubayes = posterior_mean(pstr)/.$N,
          LB.ubayes = cis[1]/.$N, 
          UB.ubayes = cis[2]/.$N
        )
      })
bayes_est <- cbind(sim_output, bayes_est)
sim_est <- left_join(sim_est, bayes_est, by = c("N", "P", "R", "CI"))

@@

  dist_L <- dist_L_gsv(heads, N, P)
  check_posterior(dist_L)
  stopifnot(length(dist_L) == heads + 1)

@@

When comparing the dishonesty rates of different groups, a possible approach is to use the
"empirical Bayes" method [@casella1985introduction]. This means that one estimates a common prior
from the pooled data, before updating the prior for each individual group.

@@

nats <- dfr %>%  
      group_by(resnat) %>% 
      summarize(
        heads = sum(heads==1), 
        N = n(),
        prop_heads = heads/N,
        prop_liars = 2*prop_heads - 1
      )
pooled_prop_liars <- mean(dfr$heads == 1) * 2 - 1

distribs <- apply(nats[, c("heads", "N")], 1, function (r) {
  N <- r["N"]
  update_prior(
      heads = r["heads"],
      N = N,
      P = 0.5, 
      prior = dbinom(0:N, N, p = pooled_prop_liars)
  )
})

@@

nat_means <- sapply(distribs, posterior_mean)
nat_cis <- sapply(distribs, posterior_quantile, q = c(0.025, 0.975))
rownames(nat_cis) <- c("ci.low", "ci.high")

@@

nats %<>% 
      mutate(
        mean = mean/N,
        ci.low = ci.low/N,
        ci.high = ci.high/N
      ) %>% 
      arrange(mean)

@@

For example, @hugh2016honesty estimates the dishonesty rates of 15 nations using a coin flip
experiment. The pooled estimate of the proportion of liars in the sample is given by *2R/N - 1* = 
`r round(pooled_prop_liars, 2)`. With *N* = `r nrow(dfr)`, this will not be much affected by 
sampling variation in the number of heads. We can use a binomial distribution with this as the prior.

@@

Results are shown in Figure \ref{fig:hj2016-pic}, with means and 95% confidence intervals. There
is substantial shrinkage from the "naïve" per-country estimates found by calculating *2R/N - 1*
separately for each country). One of the strengths of empirical Bayes,  as @casella1985introduction
points out, is that it "anticipates regression to the mean". That is, as there is noise in the data,
the most and least honest groups in the data probably had some extreme error terms. So the
British people in the sample may be less honest, and the Chinese people more honest, than the per-group 
data suggest.

@@

\[
\sum_{L = 0}^{N_{1}} \sum_{M=0}^{N_{2}} Pr(L_{1} = L) Pr(L_{2} = M) \;
  \mathbb{1}({\frac{L}{N_{1}} < \frac{M}{N_{2}})}
\]

@@

This approach has a nice interpretation, analogous to the use of the null hypothesis in frequentist
methods. We start from the belief that the proportion of liars is the same across all groups. The
posterior then shows how much the evidence changes this belief.

```{r country-comparisons}

prob_country_less <- function (ctry1, ctry2) {
  stopifnot(ctry1 %in% nats$resnat, ctry2 %in% nats$resnat)
  N_1 <- nats$N[nats$resnat == ctry1]
  N_2 <- nats$N[nats$resnat == ctry2]
  heads_1 <- nats$heads[nats$resnat == ctry1]
  heads_2 <- nats$heads[nats$resnat == ctry2]  
  diff_dist <- compare_groups(heads1 = heads_1, N1 = N_1, heads2 = heads_2, N2 = N_2, P = 0.5)

  prob_first_less <-sum(diff_dist$prob[diff_dist$diff < 0]) 
  return(prob_first_less)
}

ctry_comp <- outer(nats$resnat, nats$resnat, Vectorize(prob_country_less))
diag(ctry_comp) <- NA
rownames(ctry_comp) <- nats$resnat
colnames(ctry_comp) <- nats$resnat

selected <- c("GB", "ZA", "GR", "BR", "CN")
ctry_comp <- ctry_comp[selected, selected]
cc_hux <- as_huxtable(ctry_comp, add_colnames = TRUE, add_rownames = TRUE) %>% 
      set_caption("Bayesian country comparisons. 
            Cells give the probability that row country has fewer liars than column country.") %>% 
      set_label("tab:country-comparisons") %>% 
      set_na_string("--") %>% 
      set_number_format("%.3f") %>% 
      set_align(-1, -1, ".") %>% 
      theme_plain() %>% 
      set_bold(everywhere, 1, TRUE)

@@

cc_hux
```

# The population

So far we have been trying to estimate the proportion of *Liars* in the sample. In fact,
we usually care about the proportion of liars in populations from which the sample is drawn. This
adds another layer of randomness. When *N* is large, the sample distribution should approximate
the population distribution and either frequentist or Bayesian approaches should work. Further
simulations, not shown, suggest that Bayesian confidence intervals are less robust than frequentist ones
in bounding the population proportion of *Liars* when the sample is randomly drawn from the population.

>>>>

library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE)
options(huxtable.knitr_output_format = "latex")
source("bayesian-heads-cts.R")
suppressPackageStartupMessages({
  library(Hmisc)
  library(magrittr)
  library(dplyr)
  library(purrr)
  library(tidyr)
  library(readr)
  library(ggplot2)
  library(broom)
  library(huxtable)
  loadNamespace("MASS")
})

@@

# how many simulations to run:
nreps <- 1000L

# set to FALSE to recalculate all confidence interval estimates; TRUE for speed:
cache_results <- TRUE

# set to TRUE to (re)create the simulations CSV file, then stop:
rerun_java <- FALSE

# if TRUE, the number of liars in each simulation is randomly drawn
# from a population with `Liars` liars.
# If FALSE, the number of liars is fixed at `Liars * N`.
# If you change this, you may need to recreate the simulations CSV file
sample_from_pop <- TRUE

@@

TODO XXX: point out the claims they make. Add your sims of when the posterior is wrong
and intervals fail even for (R-T)/(N-T). Convert to sampling from a cts distrib.

@@

Some people are honest, while others are likely to lie whenever it benefits them. We would like to understand
the prevalence of lying, because dishonesty may be economically and socially harmful. Since we
cannot simply ask people if they are liars, one way to estimate the proportion of liars in a group
is to ask them to report the result of a coin flip or other random device, offering them a payment
if they report heads. Liars don't always lie: they only lie when it benefits them. So they always
report heads irrespective of the true coin flip. ^[GSV maintain this assumption and so shall I.] If
there are many more heads than we would expect by chance, we can assume many people are lying. But
how many?

@@

file_root <- paste0("GSV-sims", if (sample_from_pop) "-pop-sample")
sim_output_file <- paste0(file_root, "Output.csv")
if (rerun_java) {
  sim_input_file <- paste0(file_root, ".csv")
  readr::write_csv(sim_output, path = sim_input_file)

  # the java only lets you select a document from your home directory X-o
  copied_to_home_dir_ok <- file.copy(sim_input_file, normalizePath("~"), overwrite = TRUE)
  stopifnot(copied_to_home_dir_ok)
  stop(sprintf("Now run java -jar LyingCalculator.jar and input '%s'", sim_input_file),
        " from your home directory.\n",
        "Make sure the old output file is deleted first!\n",
        sprintf("The java will create a new file called '%s'.\n", sim_output_file), 
        "Leave it in your home directory, set `rerun_java` to FALSE and reknit this file.")
}

@@

gsv_estimates <- suppressMessages(readr::read_csv(normalizePath(file.path("~", sim_output_file))))

@@

my_prior <- dunif

bayes_est <- sim_output %>% 
      rowwise() %>% 
      do({
        pstr <- update_prior(
                heads = .$R, 
                N = .$N, 
                P = .$P/100, 
                prior = my_prior
              )
        ci <- .$CI / 100
        ci_qs <- c(1/2 - ci/2, 1/2 + ci/2)
        cis <- dist_quantile(pstr, ci_qs)
        # calculate quantities as prop dishonest
        tibble(
          EV.ubayes = dist_mean(pstr),
          LB.ubayes = cis[1], 
          UB.ubayes = cis[2]
        )
      })
bayes_est <- cbind(sim_output, bayes_est)
sim_est <- left_join(sim_est, bayes_est, by = c("N", "P", "R", "CI"))

@@

  dist_L <- dist_L_gsv(heads, N, P)
  stopifnot(length(dist_L) == heads + 1)

@@

When comparing the dishonesty rates of different groups, an interesting approach is to use the
"empirical Bayes" method [@casella1985introduction]. This piece of statistical jiu-jitsu involves
estimating a common prior from the pooled data, before updating the prior for each individual group.

@@

nats <- dfr %>%  
      group_by(resnat) %>% 
      summarize(
        heads = sum(heads==1), 
        N = n(),
        prop_heads = heads/N,
        prop_liars = 2*prop_heads - 1
      )

pp <- MASS::fitdistr(nats$prop_liars, "beta", start = list(
        shape1 = 1, shape2 = 1
      ))
pooled_prior <- function (x) dbeta(x, pp$estimate["shape1"], pp$estimate["shape2"])

distribs <- apply(nats[, c("heads", "N")], 1, function (r) {
  N <- r["N"]
  update_prior(
      heads = r["heads"],
      N = N,
      P = 0.5, 
      prior = pooled_prior
  )
})

@@

nat_means <- sapply(distribs, dist_mean)
nat_cis <- sapply(distribs, dist_quantile, probs = c(0.025, 0.975))
rownames(nat_cis) <- c("ci.low", "ci.high")

@@

nats %<>% arrange(mean)

@@

For example, @hugh2016honesty estimates the dishonesty rates of 15 nations using a coin flip
experiment. We can fit a beta distribution using the 15 observations of *2R/N - 1*. This will be
our prior.

@@

Results are shown in Figure \ref{fig:hj2016-pic}, with means and 95% confidence intervals. There is
some "shrinkage" towards the pooled mean from the naïve per-country estimates found by calculating
*2R/N - 1* separately for each country). One of the strengths of empirical Bayes,  as
@casella1985introduction points out, is that it "anticipates regression to the mean". That is, as
there is noise in the data, the most and least honest groups in the data probably had some extreme
error terms. So the British people in the sample may be less honest, and the Chinese people more
honest, than the per-group data suggest.

@@

\[
\int^1_0 \int_0^{\lambda_1} \phi(\lambda_1)\phi(\lambda_2) \; \mathrm{d}\lambda_2 \mathrm{d}\lambda_1
\]

@@

```{r country-comparisons, eval = FALSE}

prob_country_more <- function (ctry1, ctry2) {
  compare_dists(distribs[[ctry1]], distribs[[ctry2]])
}

ctry_comp <- outer(nats$resnat, nats$resnat, Vectorize(prob_country_more))
diag(ctry_comp) <- NA
rownames(ctry_comp) <- nats$resnat
colnames(ctry_comp) <- nats$resnat

selected <- c("GB", "ZA", "GR", "BR", "CN")
ctry_comp <- ctry_comp[selected, selected]
cc_hux <- as_huxtable(ctry_comp, add_colnames = TRUE, add_rownames = TRUE) %>% 
      set_caption("Bayesian country comparisons. 
            Cells give the probability that row country has higher \\lambda than column country.") %>% 
      set_label("tab:country-comparisons") %>% 
      set_na_string("--") %>% 
      set_number_format("%.3f") %>% 
      set_align(-1, -1, ".") %>% 
      theme_plain() %>% 
      set_bold(everywhere, 1, TRUE)

@@

cc_hux
